Global patterns of forest autotrophic carbon fluxes

### 2020-05-05 manuscript/appendix.Rmd
edits by krista merged

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
# knitr::opts_knit$set(root.dir = 'C:/Users/gyrcbm/Dropbox/Global_Productivity/manuscript/')
 # knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
```

>>>>

```{r setup, include = FALSE}
# knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
# knitr::opts_knit$set(root.dir = 'C:/Users/gyrcbm/Dropbox/Global_Productivity/manuscript/')
 knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
```

### 2020-05-05 manuscript/manuscript draft.Rmd
edits by krista merged

**See [issue #70](https://github.com/forc-db/Global_Productivity/issues/70), and potentially rework paragraph. **
We show that, with the exception of some differences in MAT-MAP interactions (Fig. 3), FACF are broadly consistent in their responses to climate drivers on the global scale, with no pronounced trends in C allocation among the variable pairs tested (Figs. 2, S3) . Although variation in allocation has been observed along gradients of elevation [@moser_elevation_2011] and water availability [@newman_above-_2006]--along with non-climatic axes of stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], and forest structure [@taylor_greater_2019]--little variation in relation to climate is apparent at the global scale. Our conclusion, then, is that hypothesized gradients in allocation along global climate gradients cannot currently be supported for mature forests, although data quantity and standardization is currently insufficent to rule out the possibility that such trends exist. There have been suggestions that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which are based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@delucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016], but our analysis suggests that these low values might more appropriately be attributed to the fact that these forests are old than to their tropical climate. Indeed, $CUE$ is known to decline with forest age [@delucia_forest_2007; @collalti_is_2019], but appears to be roughly independent of $GPP$ [@litton_carbon_2007]. Among our sites with relevant data, there are no signficant trends in allocation of $GPP$ to $NPP$ or of $NPP$ to $ANPP_{woody-stem}$ across gradients of latitude or climate (Fig. S3). Additional measurements with careful methodological standardization across a consistent set of mature forest sites will be necessary to determine whether any climate-driven gradients in allocation exist at the global scale.

*reworking the above paragraph:*
We show that FACF are broadly consistent in their responses to climate drivers on the global scale, with at most modest trends in C allocation among the variable pairs tested (Figs. 2, S3). 
This parallels the observation that C allocation across multiple FACF varies little with respect to climate along a steep tropical elevational gradient [@malhi_variation_2017; but see @moser_elevation_2011], 
and is not surprising given that trees face some universal constraints in carbon allocation (**REFS**). 
The one trend in allocation that emerged from our analysis was a tendency for greater C allocation aboveground--and to foliage in particular--in warm tropical climates (Fig. S3). This is consistent with observations of increasing allocation to roots with declining temperature across a tropical elevational gradient [@moser_elevation_201],   
- [@hofhansl_new_2015] (**verify**) 
- Chen et al. 2019 (doi: 10.1111/ele.13193) - find lower partitioning belowground at higher productivity (fine root NPP scales against ANPP_foliage with a slope of ~ 0.8)
- @gill_belowground_2016
- @litton_carbon_2007 - (verify)
One previously hypothesized trend that was not supported by our analysis was that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which are based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@delucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016]. Limiting the analysis to mature forests, our analysis shows no such trend (Fig. S3), suggesting that the relatively low $CUE$ of intact tropical forests is more appropriately be attributed to the fact that these forests are old than to their tropical climate. Indeed, $CUE$ is known to decline with forest age [@delucia_forest_2007; @collalti_is_2019], but appears to be roughly independent of $GPP$ [@litton_carbon_2007].

>>>>

We show that FACF are broadly consistent in their responses to climate drivers on the global scale, with at most modest trends in C allocation among the variable pairs tested (Figs. 2, S3). 
This parallels the observation that C allocation across multiple FACF varies little with respect to climate along a steep tropical elevational gradient [@malhi_variation_2017; but see @moser_elevation_2011], 
and is not surprising given that trees face some universal constraints in carbon allocation (**REFS**). 
The one trend in allocation that emerged from our analysis was a tendency for greater C allocation aboveground--and to foliage in particular--in warm tropical climates (Fig. S3). This is consistent with observations of increasing allocation to roots with declining temperature across a tropical elevational gradient [@moser_elevation_201],

- [@hofhansl_new_2015] (**verify**) 
- Chen et al. 2019 (doi: 10.1111/ele.13193) - find lower partitioning belowground at higher productivity (fine root NPP scales against ANPP_foliage with a slope of ~ 0.8)
- @gill_belowground_2016
- @litton_carbon_2007 ?

One previously hypothesized trend that was not supported by our analysis was that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which are based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@delucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016]. Limiting the analysis to mature forests, our analysis shows no such trend (Fig. S3), suggesting that the relatively low $CUE$ of intact tropical forests is more appropriately be attributed to the fact that these forests are old than to their tropical climate. Indeed, $CUE$ is known to decline with forest age [@delucia_forest_2007; @collalti_is_2019], but appears to be roughly independent of $GPP$ [@litton_carbon_2007]. (*This finding may have some important implications for modeling and our broader understanding. I'm not sure offhand how much theory has been built on the idea of low CUE in tropical forests because of warm temperatures...*)
